---
title: "qPCR analysis"
author: "Melissa Naugle"
date: "3/22/2021"
output: html_document
---

```{r setup, include=FALSE}
setwd(dir = "~/Desktop/GitHub/Symbiont_qPCR/")
rm( list = ls())
graphics.off()
library(ggplot2)
library(ggpubr)
library(plyr)
library(dplyr)
library(tidyr)
library(reshape2)
library(mosaic)
library(RColorBrewer)
library(DescTools)
library(epitools)
library(patchwork)
```

Background: These data show the ratio of species 'C' and 'D' symbionts, cell number of each symbiont, and proportion of each symbiont within corals from different sites. I want to test if symbionts are different by site (which I'm pretty sure they are based on the graph but would like to stats to show it)


# Read in data

```{r readdata}
dat1 <- read.csv("cleandata/08_05_20_fields1-4.csv")
dat2 <- read.csv("cleandata/10_14_20_fields5-8.csv")
dat3 <- read.csv("cleandata/12_14_20_redos.csv")

dat <- rbind(dat1,dat2)
nrow(dat)

#remove and replace with redos 
dat<-dat[!(dat$Sample=="Alu_F2" | dat$Sample=="TeleF5" | dat$Sample=="VatiaF6"),]

#AluF4 had only 1/3 amplify 
dat3 <- dat3[!(dat3$Sample=="AluF4"),]
dat3$Starting.Quantity..SQ. <- NULL
dat <- rbind(dat,dat3)

head(dat)
dat$FAM <- as.numeric(dat$FAM)
dat$VIC <- as.numeric(dat$VIC)
dat <- dat[!(dat$Sample=="NC"),]

dat$Site = "n"

dat <- mutate(dat, Site = derivedFactor(
  "Faga'tele" = (grepl("Tele", dat$Sample)),
  "Faga'alu" = (grepl("Alu", dat$Sample)),
  "Vatia" = (grepl("Vat", dat$Sample)),
  "Cannery" = (grepl("Can", dat$Sample)),
  "Coconut Point" = (grepl("Coco", dat$Sample)),
  .default = NA
))

dat <- dat[!is.na(dat$Site),]
```
```{r notlog}
dat_trip <- dat %>%
  group_by(Sample, Site) %>%
  summarize(FAM_mean = mean(FAM, na.rm = TRUE), VIC_mean = mean(VIC, na.rm = TRUE))

#look at data
dat_trip

#use this formula to calculate cell number for C and D
#divide by cell copy number 
dat_trip$cellnum_C <- (2^(40 - dat_trip$VIC_mean)) / 9
dat_trip$cellnum_C[is.na(dat_trip$VIC_mean)] <- (2^(40-10000000000000))
dat_trip$cellnum_D <- (2^(40 - dat_trip$FAM_mean)) / 1


#look at data
dat_trip

#calculate proportion of C and D
dat_trip$D_prop <- dat_trip$cellnum_D / (dat_trip$cellnum_C + dat_trip$cellnum_D)
dat_trip$C_prop <- dat_trip$cellnum_C / (dat_trip$cellnum_C + dat_trip$cellnum_D)

#look at data
dat_trip

#calculate C to D ratio
dat_trip$CDratio_log <- log(dat_trip$cellnum_C / dat_trip$cellnum_D)

#look at data
dat_trip

#choose bottom baseline for 100% D
dat_trip$CDratio_log[dat_trip$CDratio_log == -Inf] <- -20
dat_trip

colors = c("royalblue2", "palegreen2", "lightblue", "tomato3", "firebrick4")

#plot C:D log ratio
ggplot(dat_trip, aes(x = Site, y = CDratio_log, fill = Site)) + geom_violin() + theme_bw() + ylab("Log Cladocopium:Durisdinium Ratio") + xlab("") + geom_point(shape = 21, position = position_jitter(width = 0.001), size = 3.5, aes(fill = Site))  +theme(text=element_text(size=15), axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_text(size = 15))+ scale_x_discrete(limits = c("Faga'tele","Vatia" ,"Faga'alu","Cannery","Coconut Point" ))+ scale_color_manual(values = colors) + scale_fill_manual(values = colors)
#ggsave("figures/CDratio_2019.png", width = 7, height = 5)


dat_ratio <- dat_trip %>%
  group_by(Sample) %>%
  summarize(D_prop = (D_prop), C_prop = (C_prop))

dat_ratio <- gather(dat_ratio, "Species", "Ratio", 2:3)
dat_ratio$logRatio <- dat_ratio$Ratio

#plot props for each coral
ggplot(dat_ratio, aes(x = Sample, y = logRatio, fill = Species)) + geom_col() + theme_bw()  + ylab("Log Proportion of Durisdinium and Cladocopium") + scale_fill_manual(labels = c("Cladocopium", "Durisdinium"), values = c("violetred2", "royalblue3")) + xlab("") + theme(legend.title = element_blank()) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
#ggsave("figures/CDproportions_eachcoral_2019.png", width = 6, height = 4)

dat_ratio_means <- dat_trip %>%
  group_by(Site) %>%
  summarize(D_prop_avg = mean(D_prop), C_prop_avg = mean(C_prop))

dat_ratio_means <- gather(dat_ratio_means, "Species", "Ratio", 2:3)
dat_ratio_means$logRatio <- dat_ratio_means$Ratio


#plots mean proportions for each site
ggplot(dat_ratio_means, aes(x = Site, y = logRatio, fill = Species)) + geom_col() + theme_bw()  + ylab("Log Proportion of Durisdinium and Cladocopium") + scale_fill_manual(labels = c("Cladocopium", "Durisdinium"), values = c("violetred2", "royalblue3")) + xlab("") + theme(legend.title = element_blank())+ scale_x_discrete(limits = c("Faga'tele","Vatia" ,"Faga'alu","Cannery","Coconut Point" ))
#ggsave("figures/CDproportions_2019.png", width = 6, height = 4)


#write.csv(dat_trip, "/Users/Melissa/Desktop/GitHub/Symbiont_qPCR/cleandata/symbiont_log_ratios_07_17_21.csv", row.names = F)

```

#compare to 2014
```{r}
#2019
dat_ratio_site <- dat_trip %>%
  group_by(Site) %>%
  summarise(CD_logratio = mean(CDratio_log))
dat_ratio_site$year <- "2019"

dat_ratio_site$Site <- factor(dat_ratio_site$Site,levels = c("Faga'tele","Vatia" ,"Faga'alu","Cannery","Coconut Point"))


#2014
ratio_means_2014 <- read.csv("CRCP_2014/site_level_CDratio_2014.csv")
ratio_means_2014$year <- "2014"

ratio_means_2014$Site <- factor(ratio_means_2014$Site,levels = c("Faga'tele","Vatia" ,"Faga'alu","Cannery","Coconut Point"))

colors = c("royalblue2", "lightblue", "palegreen2", "tomato3", "firebrick4")

d2019 <- ggplot(dat_ratio_site, aes(x = Site, y = CD_logratio, fill = Site)) + geom_violin() + theme_bw() + ylab("Log Cladocopium:Durisdinium Ratio") + xlab("") + geom_point(shape = 21, size = 10, aes(fill = Site))  +theme(text=element_text(size=15), axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_text(size = 15))+ scale_color_manual(values = colors) + scale_fill_manual(values = colors) + ylim(-21,1)

d2014 <- ggplot(ratio_means_2014, aes(x = Site, y = CD_logratio, fill = Site)) + geom_violin() + theme_bw() + ylab("Log Cladocopium:Durisdinium Ratio") + xlab("") + geom_point(shape = 21, size =10, aes(fill = Site))  +theme(text=element_text(size=15), axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_text(size = 15))+ scale_color_manual(values = colors) + scale_fill_manual(values = colors)+ ylim(-21,1)

ggarrange(d2014, d2019)

#ggsave("figures/compare_2019_2014_crcp_sitelevel.png", width = 12, height = 5)


```





#I dont think this one is ideal- not meant to log transform data at beginning. Above method produces ratios in line with CRCP report 
```{r}
#try log transforming ct values first, then calc after

#dat_log <- dat
#dat_log$FAM_log <- log10(dat_log$FAM)
#dat_log$VIC_log <- log10(dat_log$VIC)


dat_ratio_log <- dat_log %>%
  group_by(Sample, Site) %>%
  summarize(FAM_log_mean = mean(FAM_log, na.rm = TRUE), VIC_log_mean = mean(VIC_log, na.rm = TRUE))
dat_ratio_log


dat_ratio_log$cellnum_C <- (2^(40 - dat_ratio_log$VIC_log_mean)) / 9
dat_ratio_log$cellnum_C[is.na(dat_ratio_log$VIC_log_mean)] <- (2^(40-10000000000000))
dat_ratio_log$cellnum_D <- (2^(40 - dat_ratio_log$FAM_log_mean)) / 1


dat_ratio_log$D_prop <- dat_ratio_log$cellnum_D / (dat_ratio_log$cellnum_C + dat_ratio_log$cellnum_D)
dat_ratio_log$C_prop <- dat_ratio_log$cellnum_C / (dat_ratio_log$cellnum_C + dat_ratio_log$cellnum_D)

dat_ratio_gather <- gather(dat_ratio_log, "Species", "Ratio", 7:8)
dat_ratio_gather$Ratio <- round(dat_ratio_gather$Ratio, digits = 4)

colors = c("royalblue2", "palegreen2", "lightblue", "tomato3", "firebrick4")

ggplot(dat_ratio_log, aes(x = Site, y = C_prop, fill = Site)) + geom_violin() + theme_bw() + ylab("Log Proportion of Cladocopium") + xlab("") + geom_point(shape = 21, position = position_jitter(width = 0.02), size = 3.5, aes(fill = Site))  + scale_x_discrete(limits = c("Faga'tele","Vatia" ,"Faga'alu","Cannery","Coconut Point" )) + scale_color_manual(values = colors) + scale_fill_manual(values = colors) +theme(text=element_text(size=15), axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_text(size = 15))

#ggsave("figures/thesis_newsymbiontfig.png", width = 6, height = 4)


#write.csv(dat_ratio_log, "/Users/Melissa/Desktop/GitHub/Symbiont_qPCR/cleandata/symbiont_log_ratios.csv", row.names = F)

dat_ratio_means <- dat_ratio_log %>%
  group_by(Site) %>%
  summarize(D_prop_avg = mean(D_prop), C_prop_avg = mean(C_prop))

dat_ratio_means <- gather(dat_ratio_means, "Species", "Ratio", 2:3)
dat_ratio_means$logRatio <- dat_ratio_means$Ratio


#plots mean proportions for each site
ggplot(dat_ratio_means, aes(x = Site, y = logRatio, fill = Species)) + geom_col() + theme_bw()  + ylab("Log Proportion of Durisdinium and Cladocopium") + scale_fill_manual(labels = c("Cladocopium", "Durisdinium"), values = c("violetred2", "royalblue3")) + xlab("") + theme(legend.title = element_blank())

#ggsave("figures/symbionts_logratio_03_25_21.pdf", width = 6, height = 4)

```


# Check assumptions 
No assumptions are met 


```{r stats}

#no assumptions met 

shapiro.test(dat_ratio$C_prop)
#not normal
#p = 2.05e-09

bartlett.test(CD_ratio ~ Site, data=dat_ratio)
#not equal
#p < 2.2e-16
```
```{r contingency}
dat_ratio_log$community_type <- "D only"

dat_ratio_log$community_type[which(dat_ratio_log$C_prop > 0)] <- "C and D" 

dat_contingency <- dat_ratio_log %>%
  select(1,2,9)


table <- table(dat_contingency$Site, dat_contingency$community_type)
table

savechisq <- chisq.test(table)
savechisq
#expected frequencies under null hypothesis
addmargins(savechisq$expected)

#since expected freqs <5 in >80% of cells, cannot run chisq or Gtest 
#GTest(dat_contingency$Site, dat_contingency$community_type)
#chisq.test(dat_contingency$Site, dat_contingency$community_type)

#can run fishers exact
fisher.test(table)
#p = 0.001327



#all show signif p values 

mosaicplot( t(table), col = c("cornflowerblue", "palegreen", "goldenrod1", "darkorange", "firebrick"), cex.axis = 1,las = 1, sub = "Symbiont Community", ylab = "Site", main = "")

#oddsratio(table, method = "wald")



```



C+D corals: 
Tele 1-6
Vatia 5-6
Alu 2





