---
title: "qPCR analysis"
author: "Melissa Naugle"
date: "3/22/2021"
output: html_document
---

```{r setup, include=FALSE}
setwd(dir = "~/Desktop/GitHub/Symbiont_qPCR/")
rm( list = ls())
graphics.off()
library(ggplot2)
library(ggpubr)
library(plyr)
library(dplyr)
library(tidyr)
library(reshape2)
library(mosaic)
library(RColorBrewer)
library(DescTools)
library(epitools)
```

Background: These data show the ratio of species 'C' and 'D' symbionts, cell number of each symbiont, and proportion of each symbiont within corals from different sites. I want to test if symbionts are different by site (which I'm pretty sure they are based on the graph but would like to stats to show it)


# Read in data

```{r readdata}
dat1 <- read.csv("cleandata/08_05_20_fields1-4.csv")
dat2 <- read.csv("cleandata/10_14_20_fields5-8.csv")
dat3 <- read.csv("cleandata/12_14_20_redos.csv")

dat <- rbind(dat1,dat2)
nrow(dat)

#remove and replace with redos 
dat<-dat[!(dat$Sample=="Alu_F2" | dat$Sample=="TeleF5" | dat$Sample=="VatiaF6"),]

dat3 <- dat3[!(dat3$Sample=="AluF4"),]
dat3$Starting.Quantity..SQ. <- NULL
dat <- rbind(dat,dat3)

head(dat)
dat$FAM <- as.numeric(dat$FAM)
dat$VIC <- as.numeric(dat$VIC)
dat <- dat[!(dat$Sample=="NC"),]

dat$Site = "n"

dat <- mutate(dat, Site = derivedFactor(
  "Faga'tele" = (grepl("Tele", dat$Sample)),
  "Faga'alu" = (grepl("Alu", dat$Sample)),
  "Vatia" = (grepl("Vat", dat$Sample)),
  "Cannery" = (grepl("Can", dat$Sample)),
  "Coconut Point" = (grepl("Coco", dat$Sample)),
  .default = NA
))

dat <- dat[!is.na(dat$Site),]
```


# Calculate ratio of C to D symbionts, cell number of each symbiont, and proportion of each symbiont

```{r calcratios}
dat_ratio <- dat %>%
  group_by(Sample, Site) %>%
  summarize(FAM_mean = mean(FAM, na.rm = TRUE), VIC_mean = mean(VIC, na.rm = TRUE))
head(dat_ratio)

#not the proper ratio but hehe oh well 
dat_ratio$CD_ratio <- (dat_ratio$FAM_mean / dat_ratio$VIC_mean)
dat_ratio$CD_ratio[is.na(dat_ratio$CD_ratio)] <- 0


#maybe look at qPCR for other spp to calc ratios 
#need to log transform 
#how did tom calc ratios - try to get within same range ?

#The proportion of clade D in each colony was then calculated as D cells/(C cells + D cells), with the number of cells of each clade calculated as 2^(40-avg. CT) divided by the gene copy numbers (Cunning and Baker 2012). from cunning 2013 pf 110
#prop of D 
#copy num = 9 and 1 for C and D


dat_ratio$cellnum_C <- (2^(40 - dat_ratio$VIC_mean)) / 9
dat_ratio$cellnum_C[is.na(dat_ratio$cellnum_C)] <- (2^(40-10000000000000))
dat_ratio$cellnum_D <- (2^(40 - dat_ratio$FAM_mean)) / 1


dat_ratio$D_prop <- dat_ratio$cellnum_D / (dat_ratio$cellnum_C + dat_ratio$cellnum_D)
dat_ratio$C_prop <- dat_ratio$cellnum_C / (dat_ratio$cellnum_C + dat_ratio$cellnum_D)

dat_ratio_gather <- gather(dat_ratio, "Species", "Ratio", 7:8)
dat_ratio_gather$Ratio <- round(dat_ratio_gather$Ratio, digits = 4)


ggplot(dat_ratio, aes(x = Site, y = C_prop, fill = Site)) + geom_boxplot() + theme_bw() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1, binwidth = 0.04) + 
  scale_fill_brewer(palette = "Set3") + ylab("Log Proportion of Cladocopium") + xlab("") + scale_y_continuous(trans = 'log10')


```

# Check assumptions 
No assumptions are met 


```{r stats}

#no assumptions met 

shapiro.test(dat_ratio$C_prop)
#not normal
#p = 2.05e-09

bartlett.test(CD_ratio ~ Site, data=dat_ratio)
#not equal
#p < 2.2e-16
```
```{r contingency}
dat_ratio$community_type <- "D only"

dat_ratio$community_type[which(dat_ratio$C_prop > 0)] <- "C and D" 

dat_contingency <- dat_ratio %>%
  select(1,2,10)


table <- table(dat_contingency$Site, dat_contingency$community_type)
table

savechisq <- chisq.test(table)
savechisq
#expected frequencies under null hypothesis
addmargins(savechisq$expected)

#since expected freqs <5 in >80% of cells, cannot run chisq or Gtest 
#GTest(dat_contingency$Site, dat_contingency$community_type)
#chisq.test(dat_contingency$Site, dat_contingency$community_type)

#can run fishers exact?
fisher.test(table)
#p = 0.001327



#all show signif p values 

mosaicplot( t(table), col = c("cornflowerblue", "palegreen", "goldenrod1", "darkorange", "firebrick"), cex.axis = 1,las = 1, sub = "Symbiont Community", ylab = "Site", main = "")

#oddsratio(table, method = "wald")



```



C+D corals: 
Tele 1-6
Vatia 5-6
Alu 2





