---
title: "qPCR analysis"
author: "Melissa Naugle"
date: "3/22/2021"
output: html_document
---

```{r setup, include=FALSE}
setwd(dir = "~/Desktop/GitHub/Symbiont_qPCR/CRCP_2014/")
rm( list = ls())
graphics.off()
library(reshape2)
library(DescTools)
library(epitools)
library(patchwork)
library(tidyverse)
library(mosaic)
```

Background: These data show the ratio of species 'C' and 'D' symbionts, cell number of each symbiont, and proportion of each symbiont within corals from different sites. I want to test if symbionts are different by site (which I'm pretty sure they are based on the graph but would like to stats to show it)


# Read in data

```{r readdata}
dat <- read.csv("TomSCLERA_SamplesAH10_to_AH101.csv")

#check that all neg controls are negative
dat[dat$Task == "NTC",]

#remove negative controls
dat<-dat[!(dat$Task=="NTC"),]

#match with site info

#parse out AH colony number
dat$Number <- parse_number(dat$Sample.Name)

#read in site info
site_info <- read.csv("../2014_data/2014 Ahyac sample info - Sheet1.csv")

dat_sites <- merge(dat, site_info, by = "Number")

#check that all numbers had a site 
dat_sites[is.na(dat_sites$Site),]

dat_sites

#take out march sample- only keep july
#dat_sites<-dat_sites[(dat_sites$Season=="July"),]

dat_sites


dat_sites$C..Mean <- NULL
dat_sites$C..SD <- NULL
dat_sites$Target.Name <- NULL
dat_sites$Baseline.End <- NULL
dat_sites$Baseline.Start <- NULL

data <- spread(dat_sites, Reporter, C.)

data$FAM[data$FAM == "Undetermined"] <- NA
data$VIC[data$VIC == "Undetermined"] <- NA

data

data$FAM <- as.numeric(data$FAM)
data$VIC <- as.numeric(data$VIC)
```
from what I can tell, C mean is the Ct mean of the 2 replicates 

I'll use this value to calculate the CD ratios 

```{r notlog}
dat_dup <- data %>%
  group_by(Number, Site, Season) %>%
  summarize(FAM_mean = mean(FAM, na.rm = TRUE), VIC_mean = mean(VIC, na.rm = TRUE))

#look at data
dat_dup
```

#####Needs revision as of 08/05/2021
```{r }
#use this formula to calculate cell number for C and D
#divide by cell copy number 
dat_dup$cellnum_C <- (2^(40 - dat_dup$VIC_mean)) / 9
dat_dup$cellnum_C[is.na(dat_dup$VIC_mean)] <- (2^(40-10000000000000))
dat_dup$cellnum_D <- (2^(40 - dat_dup$FAM_mean)) / 1


#look at data
dat_dup

#calculate proportion of C and D
dat_dup$D_prop <- dat_dup$cellnum_D / (dat_dup$cellnum_C + dat_dup$cellnum_D)
dat_dup$C_prop <- dat_dup$cellnum_C / (dat_dup$cellnum_C + dat_dup$cellnum_D)

#look at data
dat_dup

#calculate C to D ratio
dat_dup$CDratio_log <- log(dat_dup$cellnum_C / dat_dup$cellnum_D)

#look at data
dat_dup

#choose bottom baseline for 100% D
dat_dup$CDratio_log[dat_dup$CDratio_log == -Inf] <- -20
dat_dup


#write.csv(dat_dup, "/Users/Melissa/Desktop/GitHub/Symbiont_qPCR/CRCP_2014/crcp_fromtom_symbiont_log_ratios_allsites_08_05_21.csv", row.names = F)

dat_dup$Site <- factor(dat_dup$Site,levels = c("Faga'tele","Vatia" ,"Faga'alu","Cannery","Coconut Point"))


colors = c("royalblue2", "palegreen2", "lightblue", "tomato3", "firebrick4", "white")

#plot C:D log ratio
ggplot(dat_dup, aes(x = Site, y = CDratio_log, fill = Site)) + geom_violin() + theme_bw() + ylab("Log Cladocopium:Durisdinium Ratio") + xlab("") + geom_point(shape = 21, size = 3.5, aes(fill = Site))  +theme(text=element_text(size=15), axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_text(size = 15)) + scale_x_discrete(limits = c("Faga'tele","Vatia" ,"Faga'alu","Cannery","Coconut Point" ))+ scale_color_manual(values = colors) + scale_fill_manual(values = colors)
#ggsave("CDratio_2014_crcp_07_18_21.png", width = 7, height = 5)


dat_ratio <- dat_dup %>%
  group_by(Number) %>%
  summarize(D_prop = (D_prop), C_prop = (C_prop))

dat_ratio <- gather(dat_ratio, "Species", "Ratio", 2:3)
dat_ratio$logRatio <- dat_ratio$Ratio

#plot props for each coral
ggplot(dat_ratio, aes(x = Number, y = logRatio, fill = Species)) + geom_col() + theme_bw()  + ylab("Log Proportion of Durisdinium and Cladocopium") + scale_fill_manual(labels = c("Cladocopium", "Durisdinium"), values = c("violetred2", "royalblue3")) + xlab("") + theme(legend.title = element_blank()) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
#ggsave("figures/CDproportions_eachcoral_2019.png", width = 6, height = 4)

dat_ratio_means <- dat_dup %>%
  group_by(Site, Season) %>%
  summarize(D_prop_avg = mean(D_prop), C_prop_avg = mean(C_prop))

dat_ratio_means <- gather(dat_ratio_means, "Species", "Ratio", 2:3)
dat_ratio_means$logRatio <- dat_ratio_means$Ratio


#plots mean proportions for each site
ggplot(dat_ratio_means, aes(x = Site, y = logRatio, fill = Species)) + geom_col() + theme_bw()  + ylab("Log Proportion of Durisdinium and Cladocopium") + scale_fill_manual(labels = c("Cladocopium", "Durisdinium"), values = c("violetred2", "royalblue3")) + xlab("") + theme(legend.title = element_blank())+ scale_x_discrete(limits = c("Faga'tele","Vatia" ,"Faga'alu","Cannery","Coconut Point" ))
#ggsave("figures/CDproportions_2019.png", width = 6, height = 4)


#write.csv(dat_dup, "/Users/Melissa/Desktop/GitHub/Symbiont_qPCR/CRCP_2014/crcp_fromtom_symbiont_log_ratios_08_05_21.csv", row.names = F)

```

#compare to 2014
#colony level as of 7/18 
```{r}
#2019
dat_trip <- read.csv("../cleandata/symbiont_log_ratios_08_05_21_LBSP.csv")

dat_trip$season_year <- "August 2019"

dat_trip$Site <- factor(dat_trip$Site,levels = c("Faga'tele","Vatia" ,"Faga'alu","Cannery","Coconut Point"))


#2014
dat_dup <- read.csv("crcp_fromtom_symbiont_log_ratios_allsites_08_05_21_LBSP.csv")
dat_dup$season_year <- paste(dat_dup$Season, "2014")

dat_dup$Site <- factor(dat_dup$Site,levels = c("Faga'tele","Vatia" ,"Faga'alu","Cannery","Coconut Point"))
dat_dup <- dat_dup[!(is.na(dat_dup$Site)),]

dat_dup_march <- subset(dat_dup, season_year == "March 2014")
dat_dup_july <- subset(dat_dup, season_year == "July 2014")

#update these when all sites 
colors = c("royalblue2", "lightblue", "palegreen2", "tomato3", "firebrick4")
colors1 = c("lightblue", "palegreen2","firebrick4")
colors2 = c("royalblue2", "palegreen2","tomato3")

d2019 <- ggplot(dat_trip, aes(x = Site, y = CDratio_log, fill = Site)) + geom_violin() + theme_bw() + ylab("Log Cladocopium:Durisdinium Ratio") + xlab("") + geom_point(shape = 21, size = 8, aes(fill = Site))  +theme(text=element_text(size=15), axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_text(size = 15))+ scale_color_manual(values = colors) + scale_fill_manual(values = colors) + ylim(-21,21) + ggtitle("August 2019")

d2014_march <- ggplot(dat_dup_march, aes(x = Site, y = CDratio_log, fill = Site)) + geom_violin() + theme_bw() + ylab("Log Cladocopium:Durisdinium Ratio") + xlab("") + geom_point(shape = 21, size =8, aes(fill = Site))  +theme(text=element_text(size=15), axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_text(size = 15))+ scale_color_manual(values = colors2) + scale_fill_manual(values = colors2)+ ylim(-21,21)+ ggtitle("March 2014")

d2014_july <- ggplot(dat_dup_july, aes(x = Site, y = CDratio_log, fill = Site)) + geom_violin() + theme_bw() + ylab("Log Cladocopium:Durisdinium Ratio") + xlab("") + geom_point(shape = 21, size =8, aes(fill = Site))  +theme(text=element_text(size=15), axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_text(size = 15))+ scale_color_manual(values = colors1) + scale_fill_manual(values = colors1)+ ylim(-21,21)+ ggtitle("July 2014")

d2014_march + d2014_july + d2019

ggsave("compare_2019_2014_crcp_colonylevel_titles_marchincluded_new.png", width = 15, height = 5)

```



```{r}
dat_trip <- read.csv("../cleandata/symbiont_log_ratios_08_05_21_LBSP.csv")
dat_trip$season_year <- "August 2019"
dat_trip$LBSP <- factor(dat_trip$LBSP,levels = c("Low","Moderate" ,"High"))
dat_dup <- read.csv("crcp_fromtom_symbiont_log_ratios_allsites_08_05_21_LBSP.csv")
dat_dup$season_year <- paste(dat_dup$Season, "2014")
dat_dup <- dat_dup[!(is.na(dat_dup$Site)),]
dat_dup$LBSP <- factor(dat_dup$LBSP,levels = c("Low","Moderate" ,"High"))
dat_dup_march <- subset(dat_dup, season_year == "March 2014")
dat_dup_july <- subset(dat_dup, season_year == "July 2014")



d2019 <- ggplot(dat_trip, aes(x = LBSP, y = CDratio_log, fill = LBSP)) + geom_violin() + theme_bw() + ylab("Log Cladocopium:Durisdinium Ratio") + xlab("")  +theme(text=element_text(size=15), axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_text(size = 15)) + ylim(-21,21) + ggtitle("August 2019")+geom_jitter(shape = 21, size = 4, aes(fill = LBSP)) 

d2014_march <- ggplot(dat_dup_march, aes(x = LBSP, y = CDratio_log, fill = LBSP)) + geom_violin() + theme_bw() + ylab("Log Cladocopium:Durisdinium Ratio") + xlab("") + geom_jitter(shape = 21, size =4, aes(fill = LBSP))  +theme(text=element_text(size=15), axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_text(size = 15))+ ylim(-21,21)+ ggtitle("March 2014")

d2014_july <- ggplot(dat_dup_july, aes(x = LBSP, y = CDratio_log, fill = LBSP)) + geom_violin() + theme_bw() + ylab("Log Cladocopium:Durisdinium Ratio") + xlab("") + geom_jitter(shape = 21, size =4, aes(fill = LBSP))  +theme(text=element_text(size=15), axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_text(size = 15))+ ylim(-21,21)+ ggtitle("July 2014")

d2014_march + d2014_july + d2019

ggsave("compare_2019_2014_LBSP_08_05_2021.png", width = 15, height = 5)


###bin by poll level and plot 

```

